---
title: "Initial Korean covid 19"
author: "Joel Calderin"
date: "fall 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   

```{r}
str(kRdata)
```


  
# Your Question - Is dying independent of sex?

$H_0$ There is no relationship between an individuals sex and their death from COVID-19. 


$H_A$ There is a relationship between the sex of an individual and their death from COVID-19.



## Method

The response variable is categorical and the explanatory variable is also categorical. Therefore, we will be using the Cat~Cat helper file. 


## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar graph tells us that the percentage of men who died due to COVID-19 is much higher than the percentage of women who died due to COVID-19.


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
Comparing the percentage among men and women who died; nearly twice as many men died in correlation to the women. For men, 64% of them died to COVID-19, while 36% of women died to COVID-19.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In both our Chi-squared test and fisher test, we get a P-Value of less than 0.05. This means we reject the Null hypothesis. The odds ratio in the fisher test indicates that males are 2.3 times more likely to die from COVID-19 compared to females in Korea.

## Conclusions

